\section{Introduction}

In real RTL projects, a large portion of top-level and subsystem code does not describe core datapath behavior. Instead, it connects existing modules, wrappers, buses, channels, and clock/reset infrastructure. A ready/valid stream that appears as one semantic edge in a block diagram expands into \texttt{data}, \texttt{valid}, \texttt{ready}, and optional sideband signals in SystemVerilog. In AXI, TileLink, NoC, CDC, low-power, and multi-reset-domain designs, this wiring grows rapidly. For human engineers, this code is repetitive, hard to review, and easy to break. For LLMs, it is worse: the model must track long-range textual dependencies and implicit protocol constraints across a large context.

This paper takes the position that an LLM should not be asked to directly generate final RTL for integration-heavy designs. Instead, the LLM should propose candidate module connection graphs and contract drafts, while correctness is enforced by a compiler and downstream verification tools. Prior work such as VerilogEval, RTLCoder, and OpenLLM-RTL shows that RTL generation can be evaluated with automated benchmarks, but direct Verilog generation still suffers from correctness and generalization failures~\cite{verilogeval2023,rtlcoder2023,openllmrtl2025}. ChipNeMo demonstrates the value of domain adaptation for chip design tasks~\cite{chipnemo2023}. CPPL further shows that moving LLM hardware generation into a statically checked front-end DSL/JSON IR and lowering to CIRCT is a promising direction~\cite{cppl2026}. MICO focuses on a narrower target: module interconnect, adapter synthesis, and contract validation.

MICO treats module composition as a typed graph synthesis problem. The LLM acts as a proposal engine; the compiler is the authority. The compiler accepts or rejects proposed connections based on explicit interface, role, width, protocol, clock-domain, and contract rules. Accepted designs can then be lowered to SystemVerilog wrappers, SVA collateral, and eventually CIRCT HW/ESI/Verif/LTL IR.

This paper makes four contributions:

\begin{enumerate}
  \item It formulates LLM-assisted RTL integration as a typed module-composition problem rather than a direct RTL generation problem.
  \item It introduces MICO, a small language core with clock domains, interfaces, contracts, extern modules, adapters, and compose graphs.
  \item It describes a Rust compiler architecture that turns model proposals into checked IR, deterministic diagnostics, traceable SystemVerilog wrappers, and repair prompts.
  \item It defines ModuleComposeBench, an evaluation methodology for measuring first-pass composition, repair turns, unsafe rejection, and downstream RTL validation on existing RTL/IP integration tasks.
\end{enumerate}

The current artifact is intentionally narrow. It does not claim to replace general HDLs or downstream verification. Instead, it attacks the integration layer where human engineers already rely on structure, naming discipline, and review checklists, and where LLMs are most exposed to long-range textual consistency failures.
