\section{Related Work}

VerilogEval, RTLCoder, OpenLLM-RTL, CVDP, and ChipNeMo show that LLMs can be evaluated and adapted for hardware tasks~\cite{verilogeval2023,rtlcoder2023,openllmrtl2025,cvdp2025,chipnemo2023}. These works motivate better benchmarks and domain context, but the artifact emitted by a model is still often RTL text. MICO is complementary: it does not assume the model has become trustworthy. It narrows the model's output to a checked composition representation and uses compiler diagnostics plus RTL tools as the acceptance boundary.

The evaluation target is also different. Existing RTL-generation benchmarks
stress module synthesis from functional descriptions. MICO stresses integration
failure modes that appear after modules already exist: endpoint binding,
direction, width, protocol, domain, adapter, and unsafe negative cases. This
lets the paper measure not only whether a candidate wrapper compiles, but also
whether the flow rejects compositions that a permissive text-generation prompt
could accidentally wire together.

Agentic EDA systems increasingly wrap models in tool feedback loops.
Spec2RTL-Agent uses agents to process complex specifications, VeriMaAS composes
multi-agent RTL workflows with formal feedback, Dr. RTL uses EDA feedback for
RTL optimization, and HWE-Bench moves agent evaluation to repository-level
hardware bug repair~\cite{spec2rtlagent2025,verimaas2025,drrtl2026,hwebench2026}.
MICO uses the same lesson--tools must be in the loop--but applies it before RTL
repair or optimization: the proposed interconnect is a typed data structure that
can be checked, diagnosed, repaired, and compared before being lowered.

Hardware-verification benchmarks also increasingly emphasize realistic
specifications and bug-finding. AssertLLM and AssertLLM2 evaluate assertion
generation with formal and mutation-oriented criteria~\cite{assertllm2024,assertllm2_2026}.
MICO is not an assertion synthesizer. Its SVA skeletons and directed formal
monitors are acceptance evidence for a checked module-composition graph, while
full temporal verification and CDC proof remain downstream obligations.

SystemVerilog interfaces, Chisel bundles and bulk connections, Amaranth wiring, and CIRCT ESI all demonstrate that interface abstraction has practical value~\cite{chisel_interfaces,amaranth_wiring,circt_esi}. MICO differs in its objective. It targets the integration of existing RTL/IP modules, not complete hardware description. It restricts LLM output to structured AST/JSON forms, not arbitrary host-language code. It also treats clock/reset domains, protocol contracts, adapter synthesis, and formal collateral as part of the interface type system.

CPPL is the closest recent work~\cite{cppl2026}. If MICO merely claimed ``structured language plus CIRCT,'' it would overlap too much with CPPL. The boundary is narrower: CPPL targets LLM-friendly circuit generation, while MICO targets LLM-friendly module interconnect and integration. Its central objects are interfaces, clock domains, contracts, adapters, and compose graphs.

CIRCT provides a natural downstream target through HW, ESI, and Verif/LTL~\cite{circt_hw,circt_esi}. Anvil and AutoSVA motivate treating timing safety and module interaction properties as language/tool obligations rather than comments~\cite{anvil2025,autosva2021}. MICO's v0 contract layer is not a replacement for simulation, formal verification, CDC analysis, or synthesis; it makes selected obligations explicit early so a repair loop cannot silently delete them during a free-form rewrite.
