\section{MICO Language Core}

MICO v0 has five core objects: clock domains, interfaces, extern modules, adapters, and compose graphs. A connection is legal only if both instances and ports exist, the source role is an output, the destination role is an input, the interface identities match or an explicit adapter exists, the clock domains match or an explicit CDC adapter exists, and the attached contracts satisfy refinement obligations.

\begin{verbatim}
clockdom Sys(clk, rst);

interface StreamU32 @Sys {
  producer payload:u32, valid:bool;
  consumer ready:bool;
  contract stable_payload:
    valid -> stable(payload) until ready;
}

compose Top @Sys {
  inst p: Producer;
  inst f: Fifo;
  inst c: Consumer;
  connect p.tx -> f.input;
  connect f.output -> c.rx;
}
\end{verbatim}

If any condition fails, the compiler must reject the graph and produce a structured diagnostic. Such diagnostics are designed to be fed into an LLM repair loop as JSON, making repairs local and auditable rather than free-form rewrites of generated RTL.

\subsection{Static Semantics}

The v0 checker enforces a conservative subset of the intended type system. Let an endpoint be a tuple
\[
  e = (\text{instance}, \text{port}, \text{interface}, \text{role}, \text{domain}, \text{fields}, \text{contracts}).
\]
A direct connection \(e_s \rightarrow e_t\) is legal only if \(e_s\) is a producer, \(e_t\) is a consumer, the interface protocol agrees, the clock/reset domain agrees, and each required field has a compatible width and direction. If any of these predicates fail, the graph is rejected unless an explicit adapter node states how the mismatch is resolved.

\begin{table}[t]
  \centering
  \caption{Representative MICO v0 diagnostics}
  \begin{tabular}{lll}
    \toprule
    Check & Example & Repair action \\
    \midrule
    name & duplicate instance & rename or delete \\
    endpoint & unknown port & bind declared port \\
    direction & consumer to consumer & reverse edge \\
    width & u32 to u64 & insert width adapter \\
    domain & Sys to Async & insert CDC adapter \\
    contract & missing obligation & attach contract \\
    \bottomrule
  \end{tabular}
\end{table}

This design intentionally avoids silent convenience. MICO does not insert a CDC FIFO or truncating width adapter merely because doing so would make the wrapper compile. A model or user must declare the adapter explicitly, and the compiler records it in the typed IR and generated report.
