\section{LLM Collaboration Workflow}

The MICO collaboration loop has seven stages:

\begin{enumerate}
  \item Inventory modules, ports, protocols, domains, and adapters.
  \item Ask the LLM to propose interface schemas.
  \item Ask the LLM to propose a composition graph.
  \item Run compiler checks and return structured diagnostics.
  \item Ask the LLM for a minimal JSON patch.
  \item Run lint, simulation, formal, synthesis, or Vivado checks as appropriate.
  \item Produce a traceability report from compiler artifacts.
\end{enumerate}

This loop deliberately separates proposal from validation. The LLM never becomes the source of correctness. Its output is constrained to data structures that the compiler can reject, repair, and compare across model profiles.

The provider policy also separates cost from correctness. Smoke tests and bulk prompt development run with low-cost model profiles first. Higher-cost profiles are used only after objective gates pass: structured-output validity, compiler acceptance, and relevant RTL validation. This prevents a workflow in which cost escalation hides a weak prompt or an under-specified checker.

The current provider harness records each validate-only or authenticated smoke run as a redacted JSON artifact. The record includes the provider profile, model identifier, prompt SHA-256, request settings, response JSON validity, token usage when returned by the provider, optional configured cost estimate, repair-turn count, compiler-result attachment, and EDA-result attachment. It records whether an API key was present and where it was sourced, but never stores or prints the key itself. This makes LLM comparisons reproducible enough to audit without making the secret-bearing local configuration part of the artifact.

The batch runner extends this harness from a single smoke prompt to the 57-task benchmark manifest. It generates deterministic prompts for Direct Verilog, SystemVerilog-interface, MICO source, MICO JSON AST, and MICO JSON AST with compiler-feedback repair baselines. Validate-only mode plans the full low-cost matrix without provider requests, offline-fixture mode exercises compiler and open-source EDA scoring without paid calls, and execute mode uses the OpenAI-compatible SDK with response caching. Repair turns for the JSON AST baseline feed compiler diagnostics into the repair prompt and apply the returned patch to the candidate AST before rechecking. The current artifact commits the runner and smoke validation paths; it does not yet commit a full paid multi-model result matrix.

\begin{table}[t]
  \centering
  \caption{Acceptance gates for LLM-generated candidates}
  \begin{tabular}{lll}
    \toprule
    Gate & Tool & Rejects \\
    \midrule
    JSON/schema & prompt harness & malformed output \\
    Parse/check & MICO compiler & illegal graph \\
    Emit & code generator & unsupported lowering \\
    Lint/elab & Verilator/Icarus/Yosys & invalid SV \\
    Smoke proof & SymbiYosys & broken formal tool path \\
    Sim/formal & seed sim + selected formal & behavior bugs \\
    \bottomrule
  \end{tabular}
\end{table}

Diagnostics are part of the prompt interface. The repair prompt includes the original task, module inventory, interface library, current AST, and compiler diagnostics, and asks for a minimal JSON patch. The goal is to measure whether compiler feedback reduces repair turns without letting the model rewrite unrelated structure.
