\section{Artifact Reproducibility}

MICO is packaged as a reproducible artifact, not a prompt-only demonstration.
Rust, Python, open-source RTL tools, benchmark scoring, schema validation, LLM
validation, and table generation run in the repository Ubuntu 24.04 Docker EDA
environment. The host exceptions are documented: final PDF compilation uses
Windows-host \LaTeX{}, and Vivado-specific subset synthesis uses the pinned
Windows Vivado launcher.

The release workflow freezes the Docker recipe, Rust toolchain, prompts,
provider profiles, manifests, and result hashes; runs the public-development,
held-out, and supplemental realism benchmarks; runs the aggregate and paper
table generators; validates LLM configuration without exposing credentials;
records redacted authenticated provider matrices for LLM claims; runs the
Vivado subset for Xilinx evidence; and rejects staged PDFs, logs, build outputs,
local YAML files, raw provider payloads, and secret-like strings. The
deterministic, LLM, QoR, and paper-summary tables are generated artifacts, not
hand-maintained counts. The committed numeric claim map is
\texttt{docs/release\_claim\_table.md}; release builds record the corresponding
hashes in ignored \texttt{build/} artifacts and in the anonymous review bundle.

For reviewers, the one-command gate writes a release manifest, deterministic
evidence sidecar, LLM evidence sidecar, deterministic and held-out benchmark
JSON, aggregate JSON, generated table snippets, and optional Vivado hashes when
the host subset has been run. The bundle script copies these sidecars, schemas,
prompts, manifests, sanitized LLM records when present, and the paper PDF hash
into an anonymous ZIP with a SHA-256 sidecar.
