\section{Limitations and Threats to Validity}

MICO requires an interface library and adapter library, both of which need engineering effort. A type system that is too restrictive may reduce the useful search space available to the model. CIRCT lowering also introduces integration cost. Finally, ModuleComposeBench must include realistic module-composition tasks; otherwise the benchmark could collapse into a schema-generation exercise.

These risks are manageable if the project remains narrowly scoped. The first target is not a universal HDL. It is a compiler-checked interconnect layer for LLM-assisted RTL integration.

There are also evaluation threats. The 62 deterministic tasks meet the current scale threshold and include five public-development subsystem case studies plus a held-out case-study split, but they still do not by themselves support multi-model statistical claims. Verilator, Icarus, Yosys, the supported positive-task simulations, bounded SymbiYosys smoke proofs, structural and generic-mapped Yosys QoR summaries, and the minimal SymbiYosys smoke proof catch syntax, elaboration, tool-path, and simple structural issues, but they do not prove protocol correctness or timing closure for each benchmark task. The ready/valid contract checker is a v0 subset, not arbitrary LTL. Simulation now covers all positive tasks, but 20 use task-specific directed harnesses and the remaining 16 use generated ready/valid smoke harnesses derived from traceability reports. Formal smoke checks cover the 31 single-clock positive tasks, but 14 use task-specific directed monitors and the remaining 17 use generated ready/valid formal harnesses; CDC proof remains unclaimed. QoR is limited to structural area/wire counts, generic mapped-cell counts against small hand-written references, and a four-task Vivado out-of-context measurement subset; it is not routed timing closure, technology-mapped delay for the full benchmark, or board-level implementation. The authenticated low-cost LLM matrix is a negative result under the current prompt/model settings and should not be generalized to stronger models, stricter structured-output modes, or improved prompts. LLM provider results may vary across model versions and API settings; therefore prompts, model identifiers, profile settings, output hashes, compiler results, EDA results, cost metadata, and generated artifacts must be logged. The repository template intentionally leaves model price fields unset, so cost estimates require a local ignored configuration. Finally, a benchmark built only from hand-written toy modules would overstate usefulness. The full evaluation must include more realistic IP wrappers, negative safety tasks, and transparent reporting of failed repairs.

Artifact reproducibility has one practical limitation in this revision: the Docker EDA image used for Rust, Python, RTL tools, benchmark scoring, and LLM validation does not include LaTeX. The paper PDF build therefore depends on the documented Windows-host TeX Live toolchain unless LaTeX is added to the Docker image in a future artifact revision.

MICO's safety argument is consequently layered, not absolute. The compiler should prevent known illegal compositions early; RTL tools should reject invalid generated wrappers; simulation and formal harnesses should validate behavior; and human review should inspect adapter libraries, waivers, and benchmark construction.
