\documentclass[conference]{IEEEtran}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{url}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue
}

\title{MICO: Contract-Guided Module Composition for LLM-Assisted RTL Integration}

\author{
\IEEEauthorblockN{MICO Research Team}
\IEEEauthorblockA{Research Scaffold}
}

\begin{document}

\maketitle

\begin{abstract}
Large language models can generate small Verilog fragments, but real RTL integration remains brittle because correctness depends on port binding, direction, width, handshake protocol, clock/reset domains, CDC/RDC handling, wrappers, and adapter selection. This paper presents MICO, a module-interface-contract-oriented hardware front-end language and Rust compiler framework for LLM-assisted module composition. Rather than asking an LLM to emit final RTL directly, MICO restricts model output to interface schemas, module instance graphs, connection topology, adapter plans, and contract skeletons. A deterministic compiler then performs name resolution, interface compatibility checks, protocol/domain checks, adapter legality validation, and emits traceable SystemVerilog wrappers, SVA skeletons, and future CIRCT HW/ESI/Verif/LTL IR. We define the module-composition problem, the MICO language core, compiler passes, LLM repair loop, adapter discipline, and ModuleComposeBench evaluation protocol. The current open-source artifact validates the initial end-to-end path on seven seed tasks with Rust unit tests, machine-readable diagnostics, JSON IR, traceability reports, Verilator/Icarus/Yosys checks, and a minimal SymbiYosys smoke proof. It does not yet claim per-task simulation, formal proof, QoR, or multi-model pass-rate improvements. The central argument is that LLM-assisted hardware design should move from free-form RTL text generation toward compiler-checked, reproducible module composition.
\end{abstract}

\begin{IEEEkeywords}
RTL integration, large language models, hardware description languages, static checking, module composition, SystemVerilog
\end{IEEEkeywords}

\input{sections/01_introduction}
\input{sections/02_motivation}
\input{sections/03_related_work}
\input{sections/04_language_core}
\input{sections/05_rust_toolchain}
\input{sections/06_llm_workflow}
\input{sections/07_adapters}
\input{sections/08_evaluation}
\input{sections/09_artifact_reproducibility}
\input{sections/09_limitations}
\input{sections/10_conclusion}

\bibliographystyle{IEEEtran}
\bibliography{related_work}

\end{document}
